\documentclass[nooutcomes]{ximera}
%\documentclass[space,handout,nooutcomes]{ximera}

\input{../preamble.tex}

%\usepackage{tikz}


\title{Scaling in 2D}
\author{Brad Findell}
\begin{document}
\begin{abstract}
Short-answer problems about scaling in two dimensions. 
\end{abstract}
\maketitle



%\textbf{Definition.} Under a \textbf{dilation} about center $O$ and scale factor $r>0$, the image of $P$ is 
%a point $Q$ so that $Q$ lies on \wordChoice{\choice{segment}\choice[correct]{ray}\choice{line}} 
%$\answer[format=string]{OP}$ % $\overrightarrow{OP}$ 
%and $OQ=\answer[format=string]{rOP}$.  The image of $O$ is $\answer[format=string]{O}$. 

\section{Length and Area Under Scaling}
In this section, we explore what happens to length, area, and other measures under scaling.  
 
In a previous section, we defined similarity in terms of basic rigid motions and dilations, and we showed that this definition leads to well-known results about similarity, such as the AA criterion for triangle similarity and consistent ratios of lengths between and within similar figures. A key feature of this discussion was the notion of ``scale factor,'' which describes what happens to lengths under a dilation.  From the definition of a dilation, it is clear that segments on lines through the center of dilation scale by the scale factor.  We used the side-splitter theorems to show that other segments are scaled by the same scale factor.  

\begin{question}
Two students claims that a $3\times 5$ rectangle and a $4\times 6$ rectangle are similar.  

Fred says that that they are similar because the angles are the same.  Is he correct? 
$\answer[format=string]{no}$

\begin{question}
Angles are enough to determine similarity of 
\wordChoice{\choice{polygons}\choice[correct]{triangles}\choice{quadrilaterals}\choice{circles}}
but not $\answer[format=string]{quadrilaterals}$.  
Furthermore, similarity requires a consistent scale factor.  For these rectangles the height is scaled by $4/3$ whereas the base is scaled by $\answer{6/5}$.  

\begin{question}
Ned says that they are similar because you can do the same thing (i.e., add 1) to ``both sides'' of the $3\times 5$ rectangle to get the $4\times 6$ rectangle.  Is he correct? 
$\answer[format=string]{no}$

\begin{question}
Similarity requires consistent scaling, which is about $\answer[format=string]{multiplicative}$ (not additive) change.  
\end{question}
\end{question}
\end{question}
\end{question}


\begin{question}
Complete the following sentences using words such as filling, falling, covering, wrapping, hiding, surrounding, or traveling:  

Area vs. perimeter can be thought of as $\answer[format=string]{covering}$ vs. $\answer[format=string]{surrounding}$, respectively. 

Volume vs. surface area can be thought of as $\answer[format=string]{filling}$ vs. $\answer[format=string]{wrapping}$, respectively. 
\end{question}


To explore how measures of figures change under scaling and non-scaling transformations, here are some useful strategies: 
\begin{itemize}
\item Cutting the figures and rearranging the pieces.  
\item Using ``rep-tiles.'' 
\item Using known formulas for perimeters, areas, volumes, or surface areas. 
\item Approximating with segments, squares, or cubes.  
\item Shearing.
\end{itemize}

% What about angles, perimeter, surface area, weight, and temperature
% Scaling versus stretching (other examples of non-scaling)
% Aspect ratio


\begin{question}
To estimate the length of a curve, imagine approximating it with many small segments.  Now apply a similarity transformation with a scale factor of $k$.  Each segment will scale by $\answer{k}$, so the length of the curve will be $\answer{k}$ times the original length.  
\end{question}

\begin{question}
To estimate the area of a non-polygonal region, imagine covering it approximately with a grid of squares of side length $s$.   Piecing together partial squares, suppose you count $n$ squares.  Your area estimate is then $\answer{ns^2}$.  

Now apply a similarity transformation of scale factor $k$ to both the region and the grid.  Each square in the scaled grid will have 
area $\answer{(sk)^2}$, and piecing together partial squares there will be $\answer{n}$ squares.  Thus, we estimate the area of the scaled region to be $\answer{n(sk)^2}$, which is $\answer{k^2}$ times the area of the original region.  
\end{question}

\begin{question}
When $n$ copies of a plane figure can form a figure similar to the original, the figure is called a rep-$n$-tile.  Explain briefly why any parallelogram is a rep-4-tile and also a rep-9-tile.  
\begin{freeResponse}
\begin{hint}
If a parallelogram is scaled by a factor of 2, then 4 original parallelograms can make the larger parallelogram.  If a parallelogram is scaled by a factor of 3, then 9 original parallelograms can make the larger parallelogram.  (Draw pictures.)
\end{hint}
\end{freeResponse}

In general, if a parallelogram is scaled by a factor of $k$, then $\answer{k^2}$ copies of the original parallelogram can make the scaled version.  
\end{question}

\begin{question}
Use formulas to determine what happens to the perimeter and area of a rectangle when it is scaled by $k$.  

Begin with a rectangle of base $b$ and height $h$.  After scaling the rectangle by $k$, the base will be $\answer{kb}$ and the height will be $\answer{kh}$.  

The original perimeter is $\answer{2b + 2h}$.  After scaling, the perimeter will be $\answer{2bk + 2hk}$, which is precisely $\answer{k}$ times the original perimeter.  

The original area is $\answer{bh}$.  After scaling, the area will be $\answer{(kb)(kh)}$, which is precisely $\answer{k^2}$ times the original area.  
\end{question}


\begin{question}
Use formulas to determine what happens to the circumference and area of a circle when it is scaled by $k$.  

Begin with a circle of radius $r$.  After scaling the circle by $k$, its radius will be $\answer{kr}$.  

The original circumference is $\answer{2\pi r}$.  After scaling, the circumference will be $\answer{2\pi kr}$, which is precisely $\answer{k}$ times the original circumference.  

The original area is $\answer{\pi r^2}$.  After scaling, the area will be $\answer{\pi (kr)^2}$, which is precisely $\answer{k^2}$ times the original area.  
\end{question}


\end{document}
